Determinism, Complexity, and Predictability in Computer Performance

### 2014-02-13 pecomplexity.tex
editing

\section{Introduction}
%\begin{it}
%Paragraph on computer performance, including citations to Todd paper
%and summary of the results that indicate that they're deterministic
%nonlinear dynamical systems.  Given that, we should be able to
%predict.  What benefits would accrue if we could do so: power mgmt,
%end world hunger [[this is my primary goal everyday :)]], etc.
%\end{it}
Things to add to introduction
\begin{enumerate}
\item Different kinds of complexity exist in time series and this makes choosing prediction models difficult 
\subitem NOTE: RW and chaos are both complex. One is predictable and one is not.

@@

@@

\section{Modeling Computer Performance}\label{sec:compModel}

@@

% took out for space
% \subsection{Reconstructing hidden dynamics}

@@

% took out for space
% \subsection{LMA: Using dynamics in forecasting}

@@

@@

%%FIGURE INTENT
%This figure shows that entropy changs over time with the signal as well as why the regimes for \svd were chosen. 
%%%%%%%%%%%%%%%%%
\begin{figure}[htbp]
  \centering
         \caption{
[Joshua: I think adding the colored SVD trace to this would be good or putting it above this figure but need to figure how to line them up properly. Also we need to label that the numbers on the bottom of WPE are regimes not instructions...]]The weighted permutation entropy of one run of SVD. The gray bands
    are regions where the window overlaps regimes. The window size used is
    $5,000 \times 100,000$ instructions and the word length is $4$.}\label{fig:wwpe}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=0.3\textwidth]{figs/svdipcregimescolored.png}
    \caption{The instructions per cycle of \svd. Each color corresonds to the different regimes as selected by rapid shifts in WPE, as seen in Figure \ref{fig:svd_wwpe}. From left to right each change in color represents a change in regime for 6 regimes in total. }
    \label{fig:svd_ts}
  \end{subfigure}%
  %\\
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=0.3\textwidth]{figs/SVD_wwpe}
    \caption{The MASE of ARIMA vs weighted permutation entropy. }
    \label{fig:svd_wwpe}
  \end{subfigure}
  %\begin{subfigure}{0.5\textwidth}
  %  \includegraphics[width=1.0\textwidth]{figs/LMA_vs_ARIMA}
   % \caption{The MASE values for LMA against ARIMA. The dashed line is the
   % identity, delineating the traces for which either LMA or ARIMA %performed
%    better. All traces except those from {\tt SVD$_1$} lie above the line,
%    indicating that LMA is better suited prediction method for the %computer
%   traces considered.}\label{fig:lma_vs_arima}
  %\end{subfigure}
\end{figure}

 \section{Results} % (fold)
 \label{sec:results}

@@

In this section we will validate the two key findings of this work:

[[Rephrase these]]
\begin{enumerate}

@@

\item The way structure/information/complexity is processed internally by a given process plays a crucial role in predictability.
\subitem We will have shown that the existence whether linear or nonlinear is picked up on with WPE but this point gets at whether the prediction model can use the structure or not (linear can't use nonlinear structure). The right to left shifts in \col and some of the \svd regimes and the lack of shift in \gcc illustrate this nicely.

@@

%\item Introduce MASE and explain why random walk isn't *that* naive
%\item Paragraph on how to interpret MASE

%As a figure of merit we calculate the Mean Absolute Squared Error (MASE) for each forecast.

In order to analyze correctness of each prediction we split each time series into two pieces: the first 90\% referred to as the ``learning" or ``training" signal, $\{X_{i,obs}\}_{i=1}^{n}$ and the last 10\% known as the ``test" or ``correct" signal $\{c_j\}_{j=n+1}^{k+n+1}$. The learning signal is used to train an initial model (e.g., LMA or ARIMA) as described in Section \ref{sec:compModel}. The test signal is used both to assess the models forecasting accuracy and for any refitting that may be necessary. In particular, we perform $k$ 1-step predictions, after each 1-step prediction we append the training signal with the next point in the correct signal $c_j$, refit the model taking into account the new system measurement and perform another prediction. This is repeated $k$ times to obtain $\{p_j\}_{j=n+1}^{k+n+1}$.\footnote{We would like to note that this rebuilding occurs due to a problem with ARIMA models converging to a mean prediction if too long of a prediction horizon is used, this is not a handicap of either LMA or na\"ive.}

\begin{table}[htdp]
\caption{MASE distributions for 1-step predictions at a 10\% prediction horizon over 15 runs for each signal and average wpe at word length 5 and 6 for each signal.[[Joshua:Maybe delete $l=5$ as we don't use it in any figures]] }
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
                   & MASE LMA    & MASE ARIMA &MASE na\"{i}ve & $l=5$  & $l=6$ \\
\hline
\gcc                  & $ 1.5296\pm 0.0214$ & $1.8366 \pm0.0157 $ & $1.7970\pm0.0095$& $0.9510 \pm 0.0011$ & $0.9430 \pm 0.0013$ \\

>>>>

\section{Introduction}\label{sec:intro}
%\begin{it}
%Paragraph on computer performance, including citations to Todd paper
%and summary of the results that indicate that they're deterministic
%nonlinear dynamical systems.  Given that, we should be able to
%predict.  What benefits would accrue if we could do so: power mgmt,
%end world hunger [[this is my primary goal everyday :)]], etc.
%\end{it}
Things to add to introduction
\begin{enumerate}
\item Different kinds of complexity exist in time series and this makes choosing prediction models difficult 
\subitem NOTE: RW and chaos are both complex. One is predictable and one is not.

@@

\begin{figure}[htbp]
  \centering

  \begin{subfigure}{0.5\textwidth}
    \includegraphics[width=\textwidth]{figs/gccARIMAForecast}
    \caption{\gcc ARIMA }
    \label{fig:gccARIMA}
  \end{subfigure}%
    \begin{subfigure}{0.5\textwidth}
    \includegraphics[width=\textwidth]{figs/gccLMAForecast}
    \caption{\gcc LMA}
    \label{fig:gccLMA}
  \end{subfigure}
  \\
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figs/colARIMAForecast}
    \caption{\col ARIMA}
    \label{fig:colARIMA}
  \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figs/colLMAForecast}
    \caption{\col LMA}
    \label{fig:colLMA}
  \end{subfigure}
  %\begin{subfigure}{0.5\textwidth}
  %  \includegraphics[width=1.0\textwidth]{figs/LMA_vs_ARIMA}
   \caption{
For each of these, we plot the predicted value $p_n$ against the correct value $c_n$. On this type of plot a perfect prediction lies exactly on the diagonal, that is the line $p_n = c_n$, e.g., \ref{fig:colLMA} is a near perfect prediction where-as \ref{fig:gccARIMA} is a very poor prediction. }\label{fig:gcc_vs_col}  
  %\end{subfigure}
\end{figure}

@@

\section{Experimental Methods and Modelling}\label{sec:compModel}

@@

% took out for space
 \subsection{Reconstructing hidden dynamics}

@@

% took out for space
 \subsection{LMA: Using dynamics in forecasting}

@@

\subsection{Comparing Prediction Accuracy}
In order to analyze correctness of each prediction we split each time series into two pieces: the first 90\% referred to as the ``learning" or ``training" signal, $\{X_{i,obs}\}_{i=1}^{n}$ and the last 10\% known as the ``test" or ``correct" signal $\{c_j\}_{j=n+1}^{k+n+1}$. The learning signal is used to train an initial model (e.g., LMA or ARIMA) as described in Section \ref{sec:compModel}. The test signal is used both to assess the models forecasting accuracy and for any refitting that may be necessary. In particular, we perform $k$ 1-step predictions, after each 1-step prediction we append the training signal with the next point in the correct signal $c_j$, refit the model taking into account the new system measurement and perform another prediction. This is repeated $k$ times to obtain $\{p_j\}_{j=n+1}^{k+n+1}$.\footnote{We would like to note that this rebuilding occurs due to a problem with ARIMA models converging to a mean prediction if too long of a prediction horizon is used, this is not a handicap of either LMA or na\"ive.}

\begin{table}[htdp]
\caption{MASE distributions for 1-step predictions at a 10\% prediction horizon over 15 runs for each signal and average wpe at word length 5 and 6 for each signal.[[Joshua:Maybe delete $l=5$ as we don't use it in any figures, also if we leave this (before the next section we need to point forward to the next section for PE or remove it from this table]] }
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
                   & MASE LMA    & MASE ARIMA &MASE na\"{i}ve & $l=5$  & $l=6$ \\
\hline
\gcc                  & $ 1.5296\pm 0.0214$ & $1.8366 \pm0.0157 $ & $1.7970\pm0.0095$& $0.9510 \pm 0.0011$ & $0.9430 \pm 0.0013$ \\

@@

%%FIGURE INTENT
%This figure shows that entropy changs over time with the signal as well as why the regimes for \svd were chosen. 
%%%%%%%%%%%%%%%%%
\begin{figure}[htbp]
  \centering
         \caption{
[Joshua: I think adding the colored SVD trace to this would be good or putting it above this figure but need to figure how to line them up properly. Also we need to label that the numbers on the bottom of WPE are regimes not instructions...]]The weighted permutation entropy of one run of SVD. The gray bands
    are regions where the window overlaps regimes. The window size used is
    $5,000 \times 100,000$ instructions and the word length is $4$.}\label{fig:wwpe}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=0.3\textwidth]{figs/svdipcregimescolored.png}
    \caption{The instructions per cycle of \svd. Each color corresonds to the different regimes as selected by rapid shifts in WPE, as seen in Figure \ref{fig:svd_wwpe}. From left to right each change in color represents a change in regime for 6 regimes in total. }
    \label{fig:svd_ts}
  \end{subfigure}%
  %\\
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{figs/SVD_wwpe}
    \caption{The MASE of ARIMA vs weighted permutation entropy. }
    \label{fig:svd_wwpe}
  \end{subfigure}
  %\begin{subfigure}{0.5\textwidth}
  %  \includegraphics[width=1.0\textwidth]{figs/LMA_vs_ARIMA}
   % \caption{The MASE values for LMA against ARIMA. The dashed line is the
   % identity, delineating the traces for which either LMA or ARIMA %performed
%    better. All traces except those from {\tt SVD$_1$} lie above the line,
%    indicating that LMA is better suited prediction method for the %computer
%   traces considered.}\label{fig:lma_vs_arima}
  %\end{subfigure}
\end{figure}

 \section{Validation of Primary Findings} % (fold)
 \label{sec:results}

@@

As we stated in Section \ref{sec:intro} we intend to show two primary contributions of this work:

[[Joshua: Probably need to rephrase these]]
\begin{enumerate}

@@

\item The way structure/information/complexity is processed internally by a given process plays a crucial role in predictability.

\subitem We will have shown that the existence whether linear or nonlinear is picked up on with WPE but this point gets at whether the prediction model can use the structure or not (linear can't use nonlinear structure). The right to left shifts in \col and some of the \svd regimes and the lack of shift in \gcc illustrate this nicely.

@@

 In this section we will validate these two key findings.

%This portion should justify the following claim%%%%%%%%%%%%%%%%%%%%%%
%\item The existence of predictable structure in noisy real-valued time series is quantifiable by WPE and as a result WPE is correlated with prediction accuracy (MASE)  
%%%%%%%%%%%%%%%%%%%%%%%%%

%This portion should justify the following claim%%%%%%%%%%%%%%%%%%%%%%
%\item The way structure/information/complexity is processed internally by a given process plays a crucial role in predictability.
%%%%%%%%%%%%%%%%%%%%%%%%%
